### 2018-05-23 abstract.tex
Add all source files to repo. This is the first draft.

\begin{abstract}
  GEneral Matrix Multiply (GEMM) is the most fundamental
  computational kernel routine in BLAS library.
  To achieve best performance, in-memory data must be
  prefetched into fast on-chip caches ahead of use.
  Two techniques, software prefetching and data packing,
  are utilized to effectively exploit the capability of on-chip caches.
  These techniques are sufficient for lastly-recent-used (LRU) caches,
  which is often the case in traditional high-performance
  processors used in high-end servers and supercomputers.
  But in recent years the market is meeting a diversity
  in processor design, and some high-performance processors
  are equipped with shared caches with non-LRU replacement policies.
  This rises a challenge to GEMM development in multi-threaded context.
  As several threads try to load data into the shared cache simultaneously,
  inter-thread cache conflicts would increase significantly.
  We present a Shared Cache Partition (SCP) method to
  eliminate inter-thread cache conflicts in GEMM routines,
  by partitioning the shared cache into physically disjoint
  sets and assign different sets to different threads.
  We implement SCP in the OpenBLAS library and
  evaluate the performance on Phytium 2000+,
  a 64-core AArch64 processor with private LRU L1 caches
  and shared (by every 4 cores) psudo-random L2 caches.
  The results show that conflict misses on both L1 and L2
  caches are effectively reduced and GEMM performance get improved
  by $2.75\%$--$6.91\%$.
  %% FIXME BLIS? FT1500A?
\end{abstract}