### 2018-05-23 evaluation.tex

\section{Performance Evaluation}\label{sec:evaluation}

We implement SCP in the OpenBLAS~\cite{openblas} library version 0.3.0-dev,
and evaluate it on a Phytium 2000+ processor.
The Phytium 2000+ processor is an emerging high-performance
64-core processor based on ARM's AArch64 architecture.
The 64 cores are organized into 16 clusters with each
cluster containing 4 cores.
The structure of the cluster is almost the same with
the contrived 4-core processor in Figure.~\ref{fig:hierachy}
except that all 16 clusters connect to the main memory
with their L2 caches coherented by hardare.
See Table.~\ref{tab:cluster} for architectural parameters of the cores and caches.
We restrict our evaluation to DGEMM,
as in prior work~\cite{blispar,augem,poetmicro}, for two reasons.
First, the basic idea behind SCP applies to other
variants of GEMM such as SGEMM, CGEMM and ZGEMM.
Second, the LINPACK benchmark, which is used to build the
TOP500 list of world's most powerful supercomputers (FIXME citation),
relys on the DGEMM variant.

All evaluations run with the same tiling factors
$M_r$, $N_r$, $K_c$, $M_c$ and $N_c$.
The tiling factors are computed as in Section~\ref{subsec:example},
i.e. $M_r = 4$, $N_r = 8$, $K_c = 256$, $M_c = 192$, $N_c = 1024NT$.
The parallelism is controlled by two parameters,
the number of active clusters $NC$,
and the number of active threads per cluster $NT_C$.
The total number of threads is $NT = NC \cdot NT_C$.
For the Phytium 2000+ processor, the valid ranges for $NC$ and $NT_C$
are $NC \in [1, 16]$ and $NT_C \in [1, 4]$, respectively.
Generally, DGEMM performance are presented in $flops$.
%% floating-point operations per second ($flops$).
As $NT$ varies in our evaluation,
we use another metric derived from $flops$,
the average thread efficiency,
to present performance results consistently.
The average thread efficiency is a normalized value computed as
$E_{avg} = flops / (NT \cdot \widehat{flops})$
in which $\widehat{flops}$ stands for the theoretical performance peak of a single core.
%% \begin{equation*}
%%   E = flops / (NT \cdot \widehat{flops})
%% \end{equation*}

We first show how cache sharing can harm GEMM performance
in Section~\ref{subsec:drawback}.
Then Section~\ref{subsec:benefit} demonstrates how SCP can remedy
the penalty caused by cache sharing.
Quantitative analysis of the cache miss rates
is presented in Section~\ref{subsec:analysis}.
Finally, Section~\ref{subsec:privb} discusses
the privatization of shared matrix $B_2$.
%% Three versions of DGEMM are evaluated:
%% (1) Base, the baseline version,
%% (2) SCP,  the SCP version, 
%% and (3) SCP-P, the SCP version plus privatization of shared matrix $B_2$.

\subsection{Penalty of Cache Sharing}\label{subsec:drawback}

\begin{figure}
  \centering
  \subfigure[$NT=4$]{
    \label{fig:drawback-4}
    \includegraphics[width=.31\textwidth]{figures/sec1-4}
  }
  \subfigure[$NT=8$]{
    \label{fig:drawback-8}
    \includegraphics[width=.31\textwidth]{figures/sec1-8}
  }
  \\
  \subfigure[$NT=16$]{
    \label{fig:drawback-16}
    \includegraphics[width=.31\textwidth]{figures/sec1-16}
  }
  \subfigure[$NT=32,64$]{
    \label{fig:drawback-32}
    \includegraphics[width=.31\textwidth]{figures/sec1-32}
  }
  \caption{Average thread efficiency with cache sharing}
  \label{fig:drawback}
\end{figure}

As the L2 cache is shared by cores inside the cluster,
the impact of cache sharing can be demonstrated
by comparing DGEMM performance with the same $NT$
but different $NT_C$s.
Figure.~\ref{fig:drawback} shows the evaluation results.
Each curve represents a parallelism configuration.
Different subfigures are configured with different $NT$s,
and different curves in the same subfigure are
configured with different $NT_C$s.
As there is only one configuration for $NT=64$ ($NC=16$, $NT_C=4$)
and no comparation is available,
results of $NT=64$ is combined with $NT=32$ in Figure.~\ref{fig:drawback-32}.

\begin{table}
  \centering
  \caption{Average thread efficiency with cache sharing}
  \label{tab:drawback}
  \begin{tabular}{cccccc}
    \toprule
     & $NT\!\!=\!\!4$ & $NT\!\!=\!\!8$ & $NT\!\!=\!\!16$ & $NT\!\!=\!\!32$ & $NT\!\!=\!\!64$ \\
    \midrule
    $NT_C\!\!=\!\!1$ & 91.46 & 91.30 & 90.30 & -     & - \\   
    $NT_C\!\!=\!\!2$ & 90.12 & 90.25 & 89.49 & 88.26 & - \\
    $NT_C\!\!=\!\!4$ & 84.05 & 80.61 & 79.74 & 79.81 & 71.71 \\
    \bottomrule
  \end{tabular}
\end{table}

From Figure.~\ref{fig:drawback} we can see that
with all $NT$s, $NT_C=1$ achieves the best performance,
and $NT_C=2$ follows, and $NT_C=4$ comes at the lowest efficiency.
Table.~\ref{tab:drawback} summarizes the average value
of $E_{avg}$ over matrix sizes.
The table can be viewed in two directions.
Vertically, results in the same column shows how $E_{avg}$ varies with $NT_C$.
With all $NT$s, $NT_C=1$ outperforms $NT_C=4$ by a large margin.
The gap is $7.4\%$ when $NT=4$ and roughly $10\%$ when $NT>4$.
Horizontally, results in the same row reflects the scalability of
DGEMM with certain $NT_C$s.
With $NT_C=1$, the performance scales linearly with the number of threads.
With $NT_C=2$, $E_{avg}$ suffers a $1.86\%$ lose when $NT$ grows from 4 to 32.
The situation is much worse for $NT_C=4$, $E_{avg}$ drops by $14.19\%$ from
$NT=4$ to $NT=64$!

Evaluation results above show that cache sharing has a considerable
impact on DGEMM performance.

\subsection{Effectiveness of SCP}\label{subsec:benefit}

\begin{figure}
  \centering
  \subfigure[$NT=4$]{
    \label{fig:benefit-4}
    \includegraphics[width=.31\textwidth]{figures/sec2-4}
  }
  \subfigure[$NT=8$]{
    \label{fig:benefit-8}
    \includegraphics[width=.31\textwidth]{figures/sec2-8}
  }
  \subfigure[$NT=16$]{
    \label{fig:benefit-16}
    \includegraphics[width=.31\textwidth]{figures/sec2-16}
  }
  \\
  \subfigure[$NT=32$]{
    \label{fig:benefit-32}
    \includegraphics[width=.31\textwidth]{figures/sec2-32}
  }
  \subfigure[$NT=64$]{
    \label{fig:benefit-64}
    \includegraphics[width=.31\textwidth]{figures/sec2-64}
  }
  \caption{Average thread efficiency of SCP}
  \label{fig:benefit}
\end{figure}

We evaluate SCP using the same method as in Section~\ref{subsec:drawback}.
SCP is not evaluated with $NT_C=1$ because there is no need
to partition the cache if $NT_C=1$.
Figure.~\ref{fig:benefit} shows the evaluation results.
For the purpose of comparation, results without SCP
(denoted as Base in the figure) are also presented,

\begin{table}
  \centering
  \caption{Average thread efficiency win of SCP}
  \label{tab:win}
  \setlength{\tabcolsep}{3.5pt}
  \begin{tabular}{cccccc}
    \toprule
     & $NT=4$ & $NT=8$ & $NT=16$ & $NT=32$ & $NT=64$ \\
    \midrule
    $NT_C=2$ & 1.77 & 1.69 & 1.82 & 1.78 & - \\
    $NT_C=4$ & 2.75 & 4.62 & 4.41 & 3.11 & 6.91 \\
    \bottomrule
  \end{tabular}
\end{table}

%% \begin{table}
%%   \centering
%%   \caption{Average thread efficiency ($\%$) of SCP}
%%   \label{tab:benefit}
%%   \setlength{\tabcolsep}{3.5pt}
%%   \begin{tabular}{cccccc}
%%     \toprule
%%      & $NT\!\!=\!\!4$ & $NT\!\!=\!\!8$ & $NT\!\!=\!\!16$ & $NT\!\!=\!\!32$ & $NT\!\!=\!\!64$ \\
%%     \midrule
%%     $NT_C\!\!=\!\!2$ & 91.89/1.77 & 91.94/1.69 & 91.31/1.82 & 90.05/1.78 & - \\
%%     $NT_C\!\!=\!\!4$ & 86.80/2.75 & 85.22/4.62 & 84.15/4.41 & 82.92/3.11 & 78.62/6.91 \\
%%     \bottomrule
%%   \end{tabular}
%% \end{table}

Under all parallelism configurations, 
SCP performs better than the Base
consistently over all matrix sizes.
Table.~\ref{tab:win} summarizes the performance win of SCP over Base.
%% Table.~\ref{tab:benefit} summarizes the average value
%% of $E_{avg}$ in SCP evaluation.
%% There are two values in each cell, separated by a slash.
%% $E_{avg}$ stands on the left,
%% and performance win over conventional packing
%% in the corresponding parallelism configuration
%% is shown on the right for convenience.
With $NT_C=2$, SCP performs slightly better than Base
by $1.7\%$--$1.8\%$.
The performance gap becomes larger with $NT_C=4$.
ranging from $2.75\%$ to $6.91\%$.
The largest performance gain, $6.91\%$, which is a considerable improvement,
is observed under the maximum parallelism $NT=64$.
%% Looking at Table.~\ref{tab:benefit} vertically,
%% we can see that SCP still cannot fully eliminate the impact
%% of cache sharing.
%% But it do relieve the situation to a certain degree.

\subsection{Cache Miss Rate Analysis}\label{subsec:analysis}

\begin{figure}
  \centering
  \subfigure[$NT=4$]{
    \label{fig:papi-4}
    \includegraphics[width=.31\textwidth]{figures/papi-4}
  }
  \subfigure[$NT=8$]{
    \label{fig:papi-8}
    \includegraphics[width=.31\textwidth]{figures/papi-8}
  }
  \subfigure[$NT=16$]{
    \label{fig:papi-16}
    \includegraphics[width=.31\textwidth]{figures/papi-16}
  }
  \\
  \subfigure[$NT=32$]{
    \label{fig:papi-32}
    \includegraphics[width=.31\textwidth]{figures/papi-32}
  }
  \subfigure[$NT=64$]{
    \label{fig:papi-64}
    \includegraphics[width=.31\textwidth]{figures/papi-64}
  }
  \caption{Cache miss rates with $NT_C=4$}
  \label{fig:papi}
\end{figure}

To understand why SCP outperforms Base,
and whether SCP actually eliminate inter-thread conflicts,
we use the PAPI~\cite{papi} profiling tool to
analyze the cache behavior of the GEMM program.
Four cache related hardware events are counted,
which are listed and briefly described in Table.~\ref{tab:events}.
With these hardware events,
miss rates of L1 and L2 caches can be calculated with
$\frac{L1M}{L1M+L1H}$ and $\frac{L2M}{L2M+L2H}$, respectively.
The hardware events are mesured in a per-thread manner
and all presented results are the average over threads.

\begin{table}
  \centering
  \caption{Hardware events in GEMM}
  \label{tab:events}
  \begin{tabular}{lll}
    \toprule
    Name & Event & Description \\
    \midrule
    L1M & PAPI\_L1\_DCM & L1 data cache misses \\
    L1H & PAPI\_L1\_DCH & L1 data cache hits \\
    L2M & PAPI\_L2\_DCM & L2 data cache misses \\
    L2H & PAPI\_L2\_DCH & L2 data cache hits \\
    \bottomrule
  \end{tabular}
\end{table}

Figure.~\ref{fig:papi} shows the miss rates of L1 and L2 data caches.
For the sake of space, we only present results with $NT_C=4$
as SCP achieves the largest improvement with full cache sharing.
Table.~\ref{tab:papi} summarizes the average value of cache miss rates
over matrix sizes.
Under all parallelism configurations,
both L1 and L2 cache miss rates are reduced with SCP applied.
The reduced amount is roughly $13\%$ for L1 and nearly $10\%$ for L2.

For the Phytium 2000+ processor whose L1 caches are private,
only L2 caches are specially handled by SCP.
From the results we see that miss rates on both L1 and L2 caches
are reduced effectively by SCP.
While the decrease of L2 miss rate is expected,
why L1 miss rate is also affected?
The reason is that the L2 cache is inclusive.
Conflict misses on the L2 cache not only evict data out from L2,
but also invalidate the evicted data's copy in all upper level L1 caches,
thus leading to more L1 cache misses.
In general, conflict misses on an inclusive cache can affect
performance of all upper level caches.

Results in this section prove that SCP can effectively
eliminate conflict misses on shared caches.
And the benefit propagates to upper level caches, too.

\begin{table}
  \centering
  \caption{Reduction of cache miss rates with $NT_C=4$}
  \label{tab:papi}
  \setlength{\tabcolsep}{3.5pt}
  \begin{tabular}{lccccc}
    \toprule
     & $NT=4$ & $NT=8$ & $NT=16$ & $NT=32$ & $NT=64$ \\
    \midrule
    L1     & 13.87 & 13.75 & 13.54 & 13.84 & 14.21 \\
    L2     & 10.00 & 10.00 & 8.27 & 8.56 & 7.62 \\
    \bottomrule
  \end{tabular}
\end{table}
%% \begin{table}
%%   \centering
%%   \caption{Cache miss rates ($\%$) with $NT_C=4$}
%%   \label{tab:papi}
%%   \setlength{\tabcolsep}{3.5pt}
%%   \begin{tabular}{lccccc}
%%     \toprule
%%      & $NT=4$ & $NT=8$ & $NT=16$ & $NT=32$ & $NT=64$ \\
%%     \midrule
%%     L1     & 7.81 & 7.83 & 7.87 & 7.88 & 7.88 \\
%%     L1 SCP & 6.73 & 6.75 & 6.81 & 6.81 & 6.81 \\
%%     L2     & 8.78 & 8.31 & 7.13 & 7.16 & 7.10 \\
%%     L2 SCP & 7.90 & 7.49 & 6.54 & 6.55 & 6.56 \\
%%     \bottomrule
%%   \end{tabular}
%% \end{table}

\subsection{Privatization of Shared Matrix}\label{subsec:privb}
In previous evaluations, SCP only set-partitions the shared cache
for $A_2$ matrices which are private to threads.
The shared $B_2$ matrix is still packed
in conventional way-partition style.
As mentioned in Section~\ref{subsec:example},
set-partition can also be applied to $B_2$ after $B_2$ is privatized.
In this section, we will discuss and evaluate this privatization alternative.

\begin{figure}
  %% FIXME draw this figure
  \centering
  \subfigure[Conventional]{
    \label{fig:packb.conventional}
    \includegraphics[width=.50\textwidth]{figures/strategy-conventional}
  }
  \subfigure[Full privatization]{
    \label{fig:packb.full}
    \includegraphics[width=.50\textwidth]{figures/strategy-full}
  }
  \subfigure[Partial privatization]{
    \label{fig:packb.partial}
    \includegraphics[width=.50\textwidth]{figures/strategy-partial}
  }
  \caption{Packing workload ($B_2$) of thread $T_1$ with $NT_C=4$ and $NC=4$}
  %% \caption{Packing strategies of shared matrix $B_2$ with $NT_C=4$ and $NC=4$}
  \label{fig:packb}
\end{figure}

Privatization inccurs more packing overhead because
redundant packing of $B_2$ is enforced.
Figure.~\ref{fig:packb} shows three packing strategies for $B_2$
in a parallelism configuration $NT_C=4$ and $NC=4$.
The whole packing workload of $B_2$ is divided into 16 tasks
and the packing workload of thread $T_1$ is masked by z-curves. 
%% and the task assignment is annoted in Figure.~\ref{fig:packb}.
In conventional packing, each thread takes a single task.
In full privatization, all tasks are replicated to all threads.
As a result, the packing overhead grows in proportion to $NT$
in full privatization strategy, which is unacceptable
if GEMM is highly parallelized, e.g. $NT=64$.
Partial privatization in Figure.~\ref{fig:packb.partial}
offers an alternative to full privatization.
In partial privatization, privatization only occurs inside
a cluster so that tasks belong to one cluster need not to be
replicated to other clusters.
Consequently, extra packing overhead of the partial privatization
strategy is bounded by $NT_C$.
For instance, on Phytium 2000+, overhead of packing $B_2$
in partial privatization is limited to 4 times
of that in conventional packing.

Figure.~\ref{fig:privb} shows the evaluation results of the partial
privatization strategy with $NT_C=4$ and $NC=4$.
Results under other parallelism configurations are similar.
For the purpose of comparation,
SCP with and without privatization are both presented
(denoted as SCP-P and SCP in Figure.~\ref{fig:privb}, respectively).
Figure.~\ref{fig:privb.papi} shows the cache miss rates.
Privatization achieves slightly lower miss rates on both L1 and L2 caches.
Figure.~\ref{fig:privb.ate} shows the average thread efficiency.
Despite its lower cache miss rates,
privatization suffers a performance lose compared to SCP.
Why performance drops while cache miss rates get reduced?
The reason lies in the extra packing overhead introduced by privatization.
Figure.~\ref{fig:privb.breakdown} shows the occupancy ratio
of overall executing time of various overheads.
Overheads in GEMM includes
(1) packing of $A_2$, (2) packing of $B_2$,
and (3) synchronization.
While SCP-P involves roughly the same overheads in
packing $A_2$ and synchronization as SCP,
it spends much more time in packing $B_2$.
Table.~\ref{tab:breakdown} lists the breakdown of
executing time. We can see that overhead of packing $B_2$
is multiplied by a factor of $NT_C=4$,
and the total overhead increases from $2.35\%$ to $4.72\%$.

\begin{figure}
  \centering
  \subfigure[Cache miss rates]{
    \label{fig:privb.papi}
    \includegraphics[width=.31\textwidth]{figures/privb-papi}
  }
  \subfigure[Average thread efficiency]{
    \label{fig:privb.ate}
    \includegraphics[width=.31\textwidth]{figures/privb-ate}
  }
  \subfigure[Executing time breakdown]{
    \label{fig:privb.breakdown}
    \includegraphics[width=.65\textwidth]{figures/privb-breakdown}
  }
  \caption{Performance of partial privatization with $NT_C=4$ and $NC=4$}
  \label{fig:privb}
\end{figure}

\begin{table}
  \centering
  \caption{Executing time breakdown with $NT_C=4$ and $NC=4$}
  \label{tab:breakdown}
  \setlength{\tabcolsep}{3.5pt}
  \begin{tabular}{lcccc}
    \toprule
     & Sync & PackA & PackB & GEBP\\
    \midrule
    SCP   & 1.01 & 0.53 & 0.80 & 97.65 \\
    SCP-P & 1.00 & 0.52 & 3.20 & 95.28\\
    \bottomrule
  \end{tabular}
\end{table}

Results in this section demonstrats that
privatization of the shared matrix $B_2$ can further
eliminate inter-thread cache conflicts.
But the improvement in cache performance is very small because
$B_2$ only occupies a small portion of the shared L2 cache,
and this marginal improvement is not sufficient to
balance the extra packing overhead introduced
by privatization.