### 2018-05-23 related.tex

\section{Releated Work}\label{sec:related}
The idea that all level-3 BLAS operations can be built upon
a high-performance GEMM implemention was first proposed
in~\cite{gemmbased1,gemmbased2}.
The GotoBLAS library~\cite{gotoblas}, and its successor OpenBLAS~\cite{openblas},
are instantiated based on this insight.
So optimizing GEMM has always been the central task in developing
dense linear algebra software.
The GEMM optimization has two aspects.
One is developing fast computation kernels and
the other is choosing a proper overall strategy
to optimize memory access.

%% Broadly speaking, there are two aspects of this task,
%% developing fast computation kernels and optimizing memory access.
%% The former corresponds the to the kernel component and
%% the latter is the target of overall strategies.

There are several approaches for obtaining optimized GEMM kernels,
%% (1) assembly programming, (2) auto-tuning, and (3) directive-based programming,
yielding different tradeoffs between performance and portability.
In GotoBLAS~\cite{gotoblas}, OpenBLAS~\cite{openblas} and BLIS~\cite{blis},
the kernel component of GEMM are written by domain experts in assembly.
In the case of auto-tuning,
ATLAS~\cite{atlas} adopts the auto-tuning method to
automatically generate kernels with different parameters
in C and find the best-performing one by running them
on the actual computing system.
POET~\cite{poet,poetcgo,poetmicro} and AUGEM~\cite{augem} uses a 
directive-based programming approache.
Given a GEMM kernel in C, POET inserts annotations into
the C code to direct source-to-source compiler transformations.
AUGEM uses a template-based method to match
predefined patterns in the C code and transforms the matched
C code sequence into an optimized assembly code sequence.
\textsc{Poca}~\cite{poca} leverages a wide range of architecture-specific
abstractions available in the LLVM compiler infrastructure
and proposes a series of domain-specific yet architecture-independent optimizations
to obtain high-performance kernels in a portable way.

The overall strategy of GEMM mainly concerns about
tiling factors, loop orders and parallelization.
The tiling factors $M_r$, $N_r$, $M_c$, $N_c$ and $K_c$
are essential to GEMM performance.
ATLAS~\cite{atlas} relies on auto-tuning to determine optimal
values for these factors.
Analytic techniques~\cite{analytic1,analytic2,blisanalytic} are
can also be used instead of auto-tuning.
These analytic methods generally take into consideration
the cache capacity and associativity
but assume a LRU replacement policy.
As far as we know, this paper is the first to
address the problem of cache sharing with non-LRU replacement policies,
and SCP serves as the first proposal to handle this problem.
\cite{gotogemm} gives a detailed discussion of different looping orders.
GEMM are usually parallelized only on layer 3 (the M loop),
but all loops along the M and N dimensions are potentially parallelizable.
BLIS\cite{blispar} allows the users to specify a sophisticated configuration
that more than one loops of layer 1, 3, 4 and 5
can be simultaneously parallelized.
This nested parallelization can better fit complex architecture features
like multi-sockets and hyperthreading.