### 2018-05-25 conclusion.tex
Syntax fixes

\section{Conclusion}\label{sec:conclusion}
In this paper, we present a shared cache partitioning method,
SCP, to eliminate inter-thread conflicts in GEMM programs
on architectures with shared caches.
The basic idea is partition the shared cache into physically
disjoint sets and assign different sets to different threads.
Evaluation results show that SCP can effectively reduce
cache misses on the shared cache (and upper level caches),
resulting in a consideration improvement in GEMM performance
in a multi-threaded context.
In future work, we plan to extend SCP to other dense linear
algebra computing routines.