### 2018-05-26 abstract.tex
Revision from Prof. Xue

\begin{abstract}
  GEneral Matrix Multiply (GEMM) is the most fundamental
  computational kernel routine in the BLAS library.
  To achieve high performance, in-memory data must be
  prefetched into fast on-chip caches before they are used.
  Two techniques, software prefetching and data packing,
  have been used to effectively exploit the capability of on-chip lastly-recent-used (LRU) caches,
  which are popular in traditional high-performance
  processors used in high-end servers and supercomputers.
  However, the market has recently witnessed a
  new diversity in processor design, resulting in high-performance processors
  equipped with shared caches with non-LRU replacement policies.
  This poses a challenge to the development of
  high-performance GEMM in a multi-threaded context.
  As several threads try to load data into a shared cache simultaneously,
  inter-thread cache conflicts will increase significantly.
  We present a Shared Cache Partitioning (SCP) method to
  eliminate inter-thread cache conflicts in the
  GEMM routines,
  by partitioning a shared cache into physically disjoint
  sets and assigning different sets to different threads.
  We have implemented SCP in the OpenBLAS library and
  evaluated it on Phytium 2000+,
  a 64-core AArch64 processor with private LRU L1 caches
  and shared pseudo-random L2 caches (per every four cores).
  Our experimental results show that conflict misses in both L1 and L2
  caches are effectively reduced, resulting in an improvement
  of the GEMM performance by $2.75\%$--$6.91\%$.
\end{abstract}