### 2018-05-26 conclusion.tex

\section{Conclusion}\label{sec:conclusion}
In this paper, we present a shared cache partitioning method,
SCP, to reduce inter-thread conflicts in GEMM programs
on architectures with shared non-LRU caches.
The basic idea is to partition a shared cache into physically
disjoint sets and assign different sets to different threads.
Evaluation results show that SCP can effectively reduce
cache misses for a shared cache (and its
upper level caches),
resulting in a considerable improvement in GEMM performance
in a multi-threaded context.
%In future work, we plan to extend SCP to other dense linear
%algebra computing routines.