### 2018-05-27 abstract.tex
Final revision from Prof. Xue

\begin{abstract}
  GEneral Matrix Multiply (GEMM) is the most fundamental
  computational kernel routine in the BLAS library.
  To achieve high performance, in-memory data must be
  prefetched into fast on-chip caches before they are used.
  Two techniques, software prefetching and data packing,
  have been used to effectively exploit the capability of
  on-chip lastly-recent-used (LRU) caches,
  which are popular in traditional high-performance
  processors used in high-end servers and supercomputers.
  However, the market has recently witnessed a
  new diversity in processor design, resulting in high-performance processors
  equipped with shared caches with non-LRU replacement policies.
  This poses a challenge to the development of
  high-performance GEMM in a multi-threaded context.
  As several threads try to load data into a shared cache simultaneously,
  inter-thread cache conflicts will increase significantly.
  We present a Shared Cache Partitioning (SCP) method to
  eliminate inter-thread cache conflicts in the GEMM routines,
  by partitioning a shared cache into physically disjoint
  sets and assigning different sets to different threads.
  We have implemented SCP in the OpenBLAS library and
  evaluated it on Phytium 2000+,
  a 64-core AArch64 processor with private LRU L1 caches
  and shared pseudo-random L2 caches (per four-core cluster).
  Our evaluation shows that SCP has effectively reduced the
  conflict misses in both L1 and L2 caches,
  resulting in an improvement of the GEMM performance by $2.75\%$ -- $6.91\%$.
\end{abstract}