### 2018-05-27 conclusion.tex

\section{Conclusion}\label{sec:conclusion}
In this paper, we present a shared cache partitioning method,
SCP, to reduce inter-thread cache conflicts in GEMM programs
on architectures with shared non-LRU caches.
The basic idea is to partition a shared cache into physically
disjoint sets and assign different sets to different threads.
Our experimental results show that SCP can effectively reduce the inter-thread
cache misses for a shared cache (and its upper level caches),
resulting in a considerable improvement in the GEMM performance
in a multi-threaded context.
%In future work, we plan to extend SCP to other dense linear
%algebra computing routines.